---
title: "Stardew"
author: "Gina Krynski"
date: "4/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
library(tseries)
library(TSA)
library(forecast)
library(kableExtra)
library(lubridate)
```

# Stardew Valley Data

### Game Description & Data Processing
**Stardew Valley is a farming simulation role-playing game. Players create their own characters and build up their farms and the surrounding community. Also, players can invite other players online to join their farms and work together.**
```{r}
steam <- read.csv('AllSteamData.csv')
stardew_valley <- steam %>% filter(Name == 'Stardew Valley')
stardew_valley_clean <- stardew_valley %>% 
  filter(Month != "Last 30 Days") %>%
  mutate(date = my(Month)) %>%
  rename(Avg.Players = Avg..Players) %>%
  select(date, Avg.Players)

stardew_ts <- ts(stardew_valley_clean$Avg.Players, start=c(2015,12), end=c(2021,9), frequency = 12)
autoplot(stardew_ts) + labs(y="Average Monthly Players")
```


### ACF and PACF Plots
```{R}
ggAcf(stardew_ts)
ggPacf(stardew_ts)
```

**The ACF looks most similar to an AR process ACF plot. The PACF suggests thatan  AR(1) component might be useful in a model of this data.**

## Differencing for Stationarity
```{r}
first_diff <- stardew_ts %>% diff() 
first_diff %>% ggtsdisplay()
adf.test(first_diff)
```

```{r}
second_diff <- stardew_ts %>% diff() %>% diff() 
second_diff %>% ggtsdisplay()
adf.test(second_diff)
```
**The plots of the first and second difference look very similar, with a constant mean and slight non-constant variance. Because there is no significant difference between the first and second differences, we will try models with both.**


### ARIMA(1,0,0)
```{r}
fit1 <- Arima(stardew_ts, order=c(1, 0, 0))
summary(fit1)
ggAcf(residuals(fit1))
Box.test(residuals(fit1), lag=6, type="Ljung-Box", fitdf=1)

```
### ARIMA(1,1,0)
```{r}
fit2 <- Arima(stardew_ts, order=c(1, 1, 0))
summary(fit2)
ggAcf(residuals(fit2))
Box.test(residuals(fit2), lag=6, type="Ljung-Box", fitdf=1)
```
### ARIMA(1,2,0)
```{r}
fit3 <- Arima(stardew_ts, order=c(1, 2, 0))
summary(fit3)
ggAcf(residuals(fit3))
Box.test(residuals(fit3), lag=6, type="Ljung-Box", fitdf=1)
```
### ARIMA(2,1,0)
```{r}
fit4 <- Arima(stardew_ts, order=c(2, 1, 0))
summary(fit4)
ggAcf(residuals(fit4))
Box.test(residuals(fit4), lag=6, type="Ljung-Box", fitdf=1)
```
### ARIMA(2,1,1)
```{r}
fit5 <- Arima(stardew_ts, order=c(2, 1, 1))
summary(fit5)
ggAcf(residuals(fit5))
Box.test(residuals(fit5), lag=6, type="Ljung-Box", fitdf=1)
```
### ARIMA(2,1,2)
```{r}
fit6 <- Arima(stardew_ts, order=c(2, 1, 2))
summary(fit6)
ggAcf(residuals(fit6))
Box.test(residuals(fit6), lag=6, type="Ljung-Box", fitdf=1)
```
### ARIMA (1,2,2)
```{r}
fit7 <- Arima(stardew_ts, order=c(1, 2, 2))
summary(fit7)
ggAcf(residuals(fit7))
Box.test(residuals(fit7), lag=6, type="Ljung-Box", fitdf=1)
```


### ARIMA (2,2,2)
```{r}
fit8 <- Arima(stardew_ts, order=c(2, 2, 2))
summary(fit8)
ggAcf(residuals(fit8))
Box.test(residuals(fit8), lag=6, type="Ljung-Box", fitdf=1)
```
### ARIMA (1,1,1)
```{r}
fit9 <- Arima(stardew_ts, order=c(1, 1, 1))
summary(fit9)
ggAcf(residuals(fit9))
Box.test(residuals(fit9), lag=6, type="Ljung-Box", fitdf=1)
```

## Auto ARIMA Function (ARIMA(0,1,0))
```{r}
fit_auto <- auto.arima(stardew_ts)
```

### Side by Side Comparison
```{r}
Fit <- c('ARIMA(1,0,0)','ARIMA(1,1,0)','ARIMA(1,2,0)','ARIMA(2,1,0)','ARIMA(2,1,1)','ARIMA(2,1,2)','ARIMA(1,2,2)','ARIMA(2,2,2)','ARIMA(1,1,1)', 'Auto Arima')
AICc <- c(fit1$aicc, fit2$aicc, fit3$aicc, fit4$aicc, fit5$aicc, fit6$aicc, fit7$aicc, fit8$aicc, fit9$aicc, fit_auto$aicc)
fit_comparison <- data_frame(Fit, AICc)
fit_comparison %>% kbl() %>% kable_styling()
```
## Final Model ARIMA(2,2,2)
```{r}
final_fit <- fit8
autoplot(stardew_ts) + autolayer(fit8$fitted) 

```





